---
title: "chapter_3"
author: "JEFFERSON SARANSIG"
date: "2026-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
# Generate 50 random uniformly distributed numbers:
x <- runif(50)
# Check:
x
#para especificar rango de numeros
x <- runif(50, min = 2, max = 6)
head(x)
```
Plottng a histogram
```{r}
hist(x, col = 'steelblue')
```
randmly distributed data
```{r}
x <- rnorm(50)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2) #lty is type of line --dashed, lwd is width of line
```
specifying option arguments mean and sd (asignando valores a mean y sd):
```{r}
x <- rnorm(50, mean = 5, sd = 2)
```

mean and sd:
```{r}
mean(x)

```
```{r}
sd(x) 
```
quantile (where a certain percentage of your data falls below a specific point)
```{r}
quantile(x)
```
quantile 16
```{r}
quantile(x, 0.16)
```
quantile 84
```{r}
quantile(x, 0.84)
```
sanity check
```{r}
mean(x)-sd(x)


```
```{r}
mean(x)+sd(x)


```
```{r}
# 2.5th percentile:
quantile(x, 0.025)
```
```{r}
# Should correspond to M – 2 * SD:
mean(x) - 2 * sd(x)
```
```{r}
# 97.5th percentile:
quantile(x, 0.975)
```
```{r}
# Should correspond to M + 2 * SD:
mean(x) + 2 * sd(x)
```
```{r}
#20 random numbers from a normal distribution (mean 0, SD 1)-not specified, see above
hist(rnorm(n = 20)) # execute repeatedly
```
Chapter 3 exercises
```{r}
war <- read_csv('warriner_2013_emotional_valence.csv')
war
```

```{r}
range(war$Val)
```
```{r}
filter(war, Val == min(Val) | Val == max(Val))
```
```{r}
filter(war, Val %in% range(Val))
```
```{r}
arrange(war, Val) # ascending order
```
```{r}
arrange(war, desc(Val)) # descending order
```
```{r}
mean(war$Val)
```
```{r}
sd(war$Val)
```
```{r}
mean(war$Val) - sd(war$Val)
```

```{r}
mean(war$Val) + sd(war$Val)
```
```{r}
quantile(war$Val, c(0.16, 0.84))
```
```{r}
median(war$Val)
```

```{r}
quantile(war$Val, 0.5)
```
Exercises Ch.3
With the Warriner et al. (2013) data, create a ggplot2 histogram and plot the mean
as a vertical line into the plot using geom_vline() and the xintercept aesthetic
(see Chapter 2). Can you additionally add vertical dashed lines to indicate where 68%
and 95% of the data lie? (Ignore any warning messages about binwidth that may arise).

```{r}
ggplot(war, aes(x = Val)) +
  geom_histogram(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```

Exercise 2:geo_histogram > geom_density
```{r}
ggplot(war, aes(x = Val)) +
  geom_density(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```
Exercise 3
Imagine reading a research paper about a grammaticality rating study. It is noted that
the mean acceptability rating for a particular grammatical construction is 5.25 with a
standard deviation of 0.4. Assuming normality, what is the interval within which you
expect 68% of the data to lie? What about 95% of the data? Do you think the assumption
of approximate normality is reasonable in this case?
```{r}
M <- 5.25
SD <- 0.4

lower_68 <- M - SD
upper_68 <- M + SD

lower_68
upper_68
```
95%

```{r}
lower_95 <- M - 2 * SD
upper_95 <- M + 2 * SD

lower_95
upper_95

```
problem: Likert-scale data is discrete (only whole numbers are possible responses), while the normal distribution is a continuous model (mismatch)
"While the calculated 68% and 95% intervals numerically fit within a plausible Likert scale (e.g., 1–7), there are still reasons to question the assumption of approximate normality. First, Likert-scale responses are discrete rather than continuous, which conflicts with the continuous nature that a normal distribution assumes. Second, since the mean (5.25) sits closer to the upper end of a typical scale, there is less room for values to spread out above the mean than below it, which could produce a skewed rather than symmetric distribution — undermining the normality assumption even though we cannot confirm this from the summary statistics alone."

variable based on low valence and high valence and then map into box plot



