---
title: "CLASSCODE"
author: "JEFFERSON SARANSIG"
date: "2026-08-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("janitor")
library(tidyverse)
library(gridExtra)
library(janitor)
library(gridExtra)
library(stringr) #is a package for working with text/strings
library(broom)

```

# Winter CH 1; code from ch 1 of winter textbook
```{r}
numbers <- c(18, 200, 92, 14)
sum(numbers)
mean(numbers)
sd(numbers) %>% round(digits = 2) #the "digits = " se usa solo para recordarnos, en realidad se puede hacer solo round(2). "%>%" help us string operations
names <- c("a", "b", "c")
names

```


```{r}
list.files()
head(nettle) #head() shows the first few rows of a data frame
nettle <- read.csv('nettle_1999_climate.csv')
head(nettle)
tail(nettle) #head() shows the last few rows of a data frame
```


##histogram (langs+color)
```{r} 

hist(nettle$Langs) # the $ operator extracts a specific column (or list element) by name from a data frame or list.
hist(nettle$Langs, col = 'salmon')
```


```{r}
hist(nettle$Langs, col = 'steelblue')
```

```{r}
x_values <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
mean_x <- mean(x_values)
mean_x #para sacar el valor de la media
```

```{r}
x <- c(2, 3, 4, '4')
x <- as.numeric(x)
mean(x)

```

```{r}
head(nettle)
nettle[2,5]
nettle[1:4, 1:2]
nettle[nettle$Country == 'Bangladesh', 'Langs']
nettle[1:4, ]$Langs[2]
nettle[1:4, c('Country', 'Langs')]
head(nettle[,])

```

## read.csv function
```{r}
# Load data:
nettle <- read.csv('nettle_1999_climate.csv') #"read.csv" is part of Base R

# Convert data frame to tibble:
nettle <- as_tibble(nettle)
nettle
```
##read_csv function
```{r}
nettle <- read_csv('nettle_1999_climate.csv')
nettle

```

**filter con piping**

```{r} 
nettle%>%filter(Area > 6)
```
**ggplot con points**
```{r}
ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs)) # "mapping" can be omitted
```

```{r}
ggplot(nettle, aes(x = Population, y = Area)) +
  geom_point()
```



**ggplot con texto**
```{r}
ggplot(nettle, aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```


```{r}
ggsave('nettle.png', width = 8, height = 6) #to save an image
```

```{r}
# Create plots and save them in plot1 and plot2:
plot1 <- ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs))
plot2 <- ggplot(nettle,
aes(x = MGS, y = Langs, label = Country)) +
geom_text()
# Plot double plot:

grid.arrange(plot1, plot2, ncol = 2)
```


```{r}
# Plotting pipeline with %>%:
nettle %>%
filter(MGS > 8) %>%
ggplot(aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```

```{r}
#how to inspect data frame:
str(nettle) # base R inspect the dataframe
glimpse(nettle) #tidyverse inspect the dataframe

```

```{r}
#median with $
median(nettle$Population) #in this case we need $ not comma as meadian takes one specific colum from the data frame
```

#Winter Ch 2.11.3-4


```{r}
head(nettle)
filter(nettle, Country == 'Benin')

filter(nettle, Country %in% c('Benin', 'Zaire')) #checks whether each value in a vector matches any value in a set of values — it returns TRUE/FALSE for each element

select(nettle, Langs)

filter(nettle, Country == 'Benin') %>% select(Langs)
```

```{r}
filter(nettle, Country == 'Benin') %>%
select(Population:MGS)

filter(nettle, Langs > 200)

filter(nettle, Langs > 200, Population < median(Population)) #Population < median(Population) means:Keep only countries whose population is below the median population of the whole dataset



```



```{r}
#Exercises 2.11.4
# Reduce the nettle tibble to small countries:
smallcountries <- filter(nettle, Population < 4)

# Create categorical MGS variable:
nettle_MGS <- mutate(smallcountries,
MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) #we can have string things together with the "&" like, to have 3 conditions

# Group tibble for later summarizing:
nettle_MGS_grouped <- group_by(nettle_MGS, MGS_cat)

# Compute language counts for categorical MGS variable:
summarize(nettle_MGS_grouped, LangSum = sum(Langs))
```


```{r}
# My try: Can you condense all of these steps
nettle_tibble <- filter(nettle, Population < 4) %>% mutate(MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) %>% group_by(MGS_cat) %>% summarize (LangSum = sum(Langs)) 
#mutate: i am gonna create sth new

nettle_tibble
```

##case_when
```{r}


nettle %>% 
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS < 4 ~ "very dry",
      MGS < 6 ~ "dry",
      MGS < 8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile") #para ordenar las categorias
  
  ) %>% 
  group_by(MGS_cat) %>% 
  summarize (LangSum = sum(Langs))
```


```{r} 
# get descriptive statistics for the smallpopulation data
nettle %>%
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS <  4 ~ "very dry",
      MGS <  6 ~ "dry",
      MGS <  8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile")
  ) %>%
  group_by(MGS_cat) %>%
  summarize(mean = mean(Langs) %>% round(2),
            sd = sd(Langs) %>% round(2),
            min = min(Langs),
            max = max(Langs))
```
#Winter Ch 3


```{r}
# Generate 50 random uniformly distributed numbers:
x <- runif(50)
# Check:
x
#para especificar rango de numeros
x <- runif(50, min = 2, max = 6)
head(x)
```

```{r}
#Plottng a histogram
hist(x, col = 'steelblue')
```

```{r}
#randmly distributed data
x <- rnorm(50)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2) #lty is type of line --dashed, lwd is width of line
```

```{r}
#specifying option arguments mean and sd (asignando valores a mean y sd):
x <- rnorm(50, mean = 5, sd = 2)
```

###mean and sd:
```{r}
mean(x)

```
```{r}
sd(x) 
```
##quantile 
```{r}
quantile(x) #(where a certain percentage of your data falls below a specific point)
```
quantile 16
```{r}
quantile(x, 0.16)
```
quantile 84
```{r}
quantile(x, 0.84)
```
sanity check
```{r}
mean(x)-sd(x)


```
```{r}
mean(x)+sd(x)


```
```{r}
# 2.5th percentile:
quantile(x, 0.025)
```
```{r}
# Should correspond to M – 2 * SD:
mean(x) - 2 * sd(x)
```
```{r}
# 97.5th percentile:
quantile(x, 0.975)
```
```{r}
# Should correspond to M + 2 * SD:
mean(x) + 2 * sd(x)
```
```{r}
#20 random numbers from a normal distribution (mean 0, SD 1)-not specified, see above
hist(rnorm(n = 20)) # execute repeatedly
```
## class activities

```{r}
# generate and plot a normal distribution
rnorm(50) %>%
  hist() # base r

# non-base r version
rnorm(50) %>%
  as.data.frame() %>%      # make a data frame
  rename(score = 1) %>%    # rename the variable
  ggplot(aes(x = score)) + # make the plot
  geom_histogram()

# set the mean and sd
rnorm(50, 71.82, 10.42) %>% #rnorm(number of values, mean, standard deviation)
  hist() 

# with ggplot
rnorm(50, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>% #cambia el nombre de la 1era columna a score
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# with 12 participants
rnorm(12, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")
# with 1200
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# how would we screen / filter scores > 100 and < 0
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# convert to sds
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  mutate(
    score.z = scale(score) # take the z score (standardize); z-score: How far a score is from the mean, measured in standard deviations.
  ) %>%
  ggplot(aes(x = score.z)) +
  geom_histogram(color = "black", fill = "white")
```
## CH 3 exercises
```{r}
war <- read_csv('warriner_2013_emotional_valence.csv')
war
```

```{r}
range(war$Val)
```
```{r}
filter(war, Val == min(Val) | Val == max(Val))
```
```{r}
filter(war, Val %in% range(Val))
```
```{r}
arrange(war, Val) # ascending order
```
```{r}
arrange(war, desc(Val)) # descending order
```
```{r}
mean(war$Val)
```
```{r}
sd(war$Val)
```
```{r}
mean(war$Val) - sd(war$Val)
```

```{r}
mean(war$Val) + sd(war$Val)
```
```{r}
quantile(war$Val, c(0.16, 0.84))
```
```{r}
median(war$Val)
```

```{r}
quantile(war$Val, 0.5)
```
Exercises Ch.3
With the Warriner et al. (2013) data, create a ggplot2 histogram and plot the mean
as a vertical line into the plot using geom_vline() and the xintercept aesthetic
(see Chapter 2). Can you additionally add vertical dashed lines to indicate where 68%
and 95% of the data lie? (Ignore any warning messages about binwidth that may arise).

```{r}
ggplot(war, aes(x = Val)) +
  geom_histogram(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```

Exercise 2:geo_histogram > geom_density
```{r}
ggplot(war, aes(x = Val)) +
  geom_density(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```
Exercise 3
Imagine reading a research paper about a grammaticality rating study. It is noted that
the mean acceptability rating for a particular grammatical construction is 5.25 with a
standard deviation of 0.4. Assuming normality, what is the interval within which you
expect 68% of the data to lie? What about 95% of the data? Do you think the assumption
of approximate normality is reasonable in this case?
```{r}
M <- 5.25
SD <- 0.4

lower_68 <- M - SD
upper_68 <- M + SD

lower_68
upper_68
```
95%

```{r}
lower_95 <- M - 2 * SD
upper_95 <- M + 2 * SD

lower_95
upper_95

```
problem: Likert-scale data is discrete (only whole numbers are possible responses), while the normal distribution is a continuous model (mismatch)
"While the calculated 68% and 95% intervals numerically fit within a plausible Likert scale (e.g., 1–7), there are still reasons to question the assumption of approximate normality. First, Likert-scale responses are discrete rather than continuous, which conflicts with the continuous nature that a normal distribution assumes. Second, since the mean (5.25) sits closer to the upper end of a typical scale, there is less room for values to spread out above the mean than below it, which could produce a skewed rather than symmetric distribution — undermining the normality assumption even though we cannot confirm this from the summary statistics alone."

variable based on low valence and high valence and then map into box plot


### boxplots

```{r}
# valence of words a vs p

war %>%
  mutate(
    word_type = case_when(
      str_detect(Word, "^a") ~ "a", #str_detect busca patterns within a text; ^ significa If it starts with a, put "a" in word_type
      str_detect(Word, "^p") ~ "p", #~"p" significa If the condition on the left is TRUE, put "p" in the new column
      .default = "error" #esta linea dice que hacer en caso de que no haya a ni p
    )
  ) %>%
  filter(word_type != "error") %>% #Keep only words that are not labeled "error"
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()

# creating a valence factor with levels low, medium, high
war %>%
  mutate(
    word_type = case_when(
      Val < 3 ~ "low",
      Val < 7 ~ "medium",
      Val >= 7 ~ "high",
      .default = "error"),
    word_type = fct_relevel(word_type, "low", "medium", "high")
  ) %>%
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()
```

#Winter Ch 5

```{r}
RTs <- c(600, 650, 700, 1000, 4000)
RTs
logRTs <- log(RTs)
logRTs
```

```{r}
exp(logRTs) # undo the logging
```

```{r}
ELP <- read_csv('ELP_frequency.csv')
ELP
```

```{r}
ELP <- mutate(ELP,
Log10Freq = log10(Freq),
LogRT = log(RT))
ELP
```

```{r}
#Raw frequencies
ELP %>% ggplot(aes(x = Freq, y = LogRT, label = Word)) +
geom_text() +
geom_smooth(method = 'lm') +
ggtitle('Log RT ~ raw frequency') +
theme_minimal()
```

```{r}
#log frequency values
ELP %>% ggplot(aes(x = Log10Freq, y = LogRT,
label = Word)) +
geom_text() +
geom_smooth(method = 'lm') +
ggtitle('Log RT ~ log frequency') +
theme_minimal()
```
```{r}
ELP_mdl <- lm(LogRT ~ Log10Freq, data = ELP)
tidy(ELP_mdl)
```

```{r}
b0 <- tidy(ELP_mdl)$estimate[1] # intercept
b1 <- tidy(ELP_mdl)$estimate[2] # slope
logRT_10freq <- b0 + b1 * 1
logRT_1000freq <- b0 + b1 * 3
logRT_10freq #predicted log response durations
logRT_1000freq
exp(logRT_10freq) #back to raw response durations, in miliseconds
exp(logRT_1000freq)
```
```{r}
ELP <- mutate(ELP,
Log10Freq_c = Log10Freq - mean(Log10Freq),
Log10Freq_z = Log10Freq_c / sd(Log10Freq_c)) #how many standard deviations a word's Log10Freq is from the mean
select(ELP, Freq, Log10Freq, Log10Freq_c, Log10Freq_z)
```

```{r}
# Same as before, different approach:
ELP <- mutate(ELP,
Log10Freq_c = scale(Log10Freq, scale = FALSE),
Log10Freq_z = scale(Log10Freq))
```

linear models with the centered and standardized
predictors:
```{r}

ELP_mdl_c <- lm(LogRT ~ Log10Freq_c, ELP) # centered
ELP_mdl_z <- lm(LogRT ~ Log10Freq_z, ELP) # z-scored
```
Comparing the coefficients from the different models:
```{r}
tidy(ELP_mdl) %>% select(term, estimate)
```
```{r}
tidy(ELP_mdl_c) %>% select(term, estimate) # x = 0 represents the mean of Log10Freq;the intercept (6.507966) is the predicted response at the mean; slope: a 1-unit increase in Log10Freq = 0.104-unit decrease in response
```

```{r}
tidy(ELP_mdl_z) %>% select(term, estimate) # Slope changes: 1 unit of Log10Freq_z = 1 SD of Log10Freq. For a 1 SD increase in Log10Freq, the predicted response decreases by 0.150103 units
```
```{r}
glance(ELP_mdl)$r.squared
glance(ELP_mdl_c)$r.squared
glance(ELP_mdl_z)$r.squared #r.squared: how much of the variation in Y is explained by the regression model
```

```{r}
glance(lm(LogRT ~ Freq, ELP))$r.squared # Nonlinear transformations of X can change R² and model fit
```
```{r}
with(ELP, cor(Log10Freq, LogRT)) #correlation is negative means that, as frequency increases, response durations become shorter.
```
## Recreates correlation using regression with z-scored X and Y; no intercept needed
```{r}
ELP_cor <- lm(scale(LogRT) ~ -1 + Log10Freq_z, ELP)
tidy(ELP_cor) %>% select(estimate) #explanation on p101
```
##exercise ch5-guesstimate
#Problem set 1
```{r}
data_stress <- read_csv("PS1/data_stress.csv")
data_answers <- read_csv("PS1/data_answers.csv")
glimpse(data_stress)

```
```{r}
glimpse(data_answers)
```

```{r}
names(data_stress)
names(data_answers)
```

```{r}
unique(data_stress$screen) # NA "instructions" "countdown"    "stimulus"     "response"    "timeout" 
  data_stress_clean <- filter(data_stress, screen == "response") #screen column has the participants' responses
data_stress_clean 
```

```{r}
nrow(data_stress_clean) #the number of rows is the same as in the instructions
```
```{r}
length(unique(data_stress_clean$id)) #it keeps the 38 participants
count(data_stress_clean, id) #all 38 participants have 80 trials
```

```{r}
count(data_stress_clean, type) #matches 38x48 (distractor) and 38x32(target_items)
```
##cleaning
In the data_stress 'el' has an accent, while in the answer sheet it does not. Also, what is ´eu´ in the answer sheet? typo?
```{r}
unique(data_stress_clean$response) #particiapnts response: "tu"       NA         "yo"       "él"       "ellos"    "nosotros"
unique(data_answers$answer) #answer sheet: "nosotros" "yo"       "ellos"    "tu"       "el"       "eu" 
```
```{r}
filter(data_answers, answer == "eu") #so I am assuming there is a typo in the answer column and i will change eu into tu
```
##Changing the answers
```{r}
data_stress_clean$response <- ifelse(data_stress_clean$response == "él", "el", data_stress_clean$response) #I am changing the el with accent with the el without accent and without modifying the rest
unique(data_stress_clean$response) #[1] "tu"       NA         "yo"       "el"       "ellos"    "nosotros"

sum(is.na(data_stress_clean$response)) #????
```
```{r}
data_answers$answer <- ifelse(data_answers$answer == "eu", "tu", data_answers$answer)
unique(data_answers$answer)
```
```{r}
any(duplicated(data_answers$sentence)) #[1] TRUE
nrow(data_answers) #[1] 130
```
```{r}
data_answers[duplicated(data_answers$sentence), ]
```

```{r}
print(data_answers[duplicated(data_answers$sentence), c("sentence", "verb conjugation", "answer")], width = Inf)
```
```{r}
subset(data_answers, sentence == "Aca lo invito")
```

```{r}
head(data_stress_clean)
head(data_answers)
unique(data_answers$answer)
```


```{r}
data_stress_clean
head(data_stress_clean[, c("verb", "tense", "audio")], 12)
unique(data_stress_clean$audio)
```


```{r}
head(unique(data_stress_clean$verb), 20)
head(unique(data_answers$`verb conjugation`), 20)
```
##filenames extraction
Extracting sentences from audio column in data_stress
```{r}
data_stress_clean <- data_stress_clean %>%
  mutate(sentence_clean = gsub(".*_", "", audio)) %>% # keep only what comes after the last underscore
  mutate(sentence_clean = gsub(".mp3", "", sentence_clean)) # remove the ".mp3"
head(data_stress_clean$sentence_clean)
```
```{r}
setdiff(data_stress_clean$sentence_clean, data_answers$sentence) #character(0)
```
```{r}
any(duplicated(paste(data_answers$sentence, data_answers$tense))) #paste() glues the two columns into one text per row, so this asks, "does any sentence + tense combination appear more than once?"
```
```{r}
unique(data_stress_clean$tense)
unique(data_answers$tense) #tenses are written similarly in both files
```
```{r}
head(data_stress_clean)
head(data_answers_clean)
```
##merged
```{r}
data_answers_clean <- data_answers[, c("sentence", "tense", "answer")]
data_stress_merged <- merge(data_stress_clean, data_answers_clean,
                            by.x = c("sentence_clean", "tense"),
                            by.y = c("sentence", "tense")
                            )

```
##scored

```{r}
data_stress_scored <- data_stress_merged %>%
  mutate(score = ifelse(response == answer, 1, 0))

data_stress_scored %>%
  select(response, answer, score) %>%
  head(10)
```


```{r}
head(data_stress_scored)
```
##mean participant/type
```{r}
data_stress_means <- data_stress_scored %>%
  group_by(id, type) %>% #each combination of participant (id) and type is a  group
  summarise(mean_score = mean(score, na.rm = TRUE)) #mean_score is a new colum w/ each group's mean
head(data_stress_means)
unique(data_stress_scored$response) #this should always keep el, tu without the accent
```
##data proficiency
```{r}
data_proficiency <- read_csv("PS1/data_proficiency.csv")
glimpse(data_proficiency)
names(data_proficiency) # [1] "id"  "trial"   "response"  [4] "onset_time"             "clock_time"             "reaction_time"         [7] "absolute_onset_time"    "absolute_clock_time"    "absolute_reaction_time" [10] "tag"                    "answer"   
dim(data_proficiency) #[1] 3188   11
```
```{r}
unique(data_proficiency$tag) #tag tiene NA        "rawr"    "click"   "timeout"
data_proficiency %>%
  select(id, trial, response, tag, answer) %>%
  head(20)

```
```{r}
count(data_proficiency, tag) #dentro de column tag, click parece ser la respuesta del participant
count(data_proficiency, id) #la mayoria tiene 85 trials, excepto 2
data_proficiency_counts <- count(data_proficiency, id)
subset(data_proficiency_counts, n != 85) #sale los datos de los 2 que no completaron todos los ejercicios. Decidi no incluirlos
```
```{r}
subset(data_proficiency, id == 15423803) #BORRAR
unique(data_proficiency$tag)
```


```{r}
table(data_proficiency$tag, is.na(data_proficiency$response)) #is.na asks is `response' missing, FALSE=not missing
head(subset(data_proficiency, tag == "click"), 10) #this shows that the tag "click" has participants' answers
head(subset(data_proficiency, tag == "rawr"), 10) #response have NA
head(subset(data_proficiency, tag == "timeout"), 10) #response have literally "timeout"
```
###timeout into NA
```{r}
data_proficiency_clean <- data_proficiency %>%
  filter(tag == "click" | tag == "timeout") %>%
  mutate(response = ifelse(response == "timeout", NA, response))
unique(data_proficiency_clean$response) #ahora los "timeout" cambiaron a NA en al columna response
nrow(data_proficiency_clean) #1862, the same as above with table(it was just to check)
sum(is.na(data_proficiency_clean$response)) #to check if "timeout" change into NA
```
```{r}
data_proficiency_clean <- data_proficiency_clean %>%
  select(id, trial, response, answer)
head(data_proficiency_clean)
data_proficiency_scored <- data_proficiency_clean %>%
  mutate(score = ifelse(response == answer, 1, 0))
table(data_proficiency_scored$score, useNA = "ifany")
```

```{r}
head(subset(data_proficiency_scored, score == 0, select = c(response, answer)), 20)
```
```{r}
setdiff(data_proficiency_scored$response, data_proficiency_scored$answer)
```
```{r}
setdiff(data_proficiency_scored$answer, data_proficiency_scored$response) #this shows that all answers where chosen at least once
head(data_proficiency_scored)

data_proficiency_means <- data_proficiency_scored %>%
  group_by(id) %>% #solo id porque habian mas de una columna por id y ahora solo 1
  summarise(proficiency = mean(score, na.rm = TRUE))
nrow(data_proficiency_means) #38 en total, that is, #of participants
head(data_proficiency_means)
```
```{r}
count(data_proficiency_clean, id) %>% filter(n != 50)
incomplete_ids <- c(15411133, 15423803)
incomplete_ids
data_proficiency_means <- data_proficiency_means %>%
  filter(!(id %in% incomplete_ids)) #significa: manten todos exepto los de arriba
nrow(data_proficiency_means) #36, ya excluidos los dos de arriba
```
##incomplete_ids
```{r}
data_stress_means <- data_stress_means %>%
  filter(!(id %in% incomplete_ids))
```
###ungrouping incomplete ids
```{r}
#ungrouping the two participants who did not finish the task
data_stress_means <- data_stress_means %>%
  ungroup() %>%
  filter(!(id %in% incomplete_ids))
nrow(data_stress_means) #72, that is, 36 participants x 2types
```
```{r}
data_stress_wide <- data_stress_means %>%
  pivot_wider(names_from = type, values_from = mean_score) #makes one extra column instead of 2 rows
head(data_stress_wide)
nrow(data_stress_wide)
```

```{r}
data_final <- merge(data_stress_wide, data_proficiency_means, by = "id")
head(data_final)
nrow(data_final)
sum(is.na(data_final)) #sanity check, debe ser 0
```

