---
title: "CLASSCODE"
author: "JEFFERSON SARANSIG"
date: "2026-08-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("janitor")
library(tidyverse)
library(gridExtra)
library(janitor)
library(gridExtra)
library(stringr) #is a package for working with text/strings
library(broom)

```

# Winter CH 1; code from ch 1 of winter textbook
```{r}
numbers <- c(18, 200, 92, 14)
sum(numbers)
mean(numbers)
sd(numbers) %>% round(digits = 2) #the "digits = " se usa solo para recordarnos, en realidad se puede hacer solo round(2). "%>%" help us string operations
names <- c("a", "b", "c")
names

```


```{r}
list.files()
head(nettle) #head() shows the first few rows of a data frame
nettle <- read.csv('nettle_1999_climate.csv')
head(nettle)
tail(nettle) #head() shows the last few rows of a data frame
```


##histogram (langs+color)
```{r} 

hist(nettle$Langs) # the $ operator extracts a specific column (or list element) by name from a data frame or list.
hist(nettle$Langs, col = 'salmon')
```


```{r}
hist(nettle$Langs, col = 'steelblue')
```

```{r}
x_values <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
mean_x <- mean(x_values)
mean_x #para sacar el valor de la media
```

```{r}
x <- c(2, 3, 4, '4')
x <- as.numeric(x)
mean(x)

```

```{r}
head(nettle)
nettle[2,5]
nettle[1:4, 1:2]
nettle[nettle$Country == 'Bangladesh', 'Langs']
nettle[1:4, ]$Langs[2]
nettle[1:4, c('Country', 'Langs')]
head(nettle[,])

```

## read.csv function
```{r}
# Load data:
nettle <- read.csv('nettle_1999_climate.csv') #"read.csv" is part of Base R

# Convert data frame to tibble:
nettle <- as_tibble(nettle)
nettle
```
##read_csv function
```{r}
nettle <- read_csv('nettle_1999_climate.csv')
nettle


```

**filter con piping**

```{r} 
nettle%>%filter(Area > 6)
```
**ggplot con points**
```{r}
ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs)) # "mapping" can be omitted
```

```{r}
ggplot(nettle, aes(x = Population, y = Area)) +
  geom_point()
```



**ggplot con texto**
```{r}
ggplot(nettle, aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```


```{r}
ggsave('nettle.png', width = 8, height = 6) #to save an image
```

```{r}
# Create plots and save them in plot1 and plot2:
plot1 <- ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs))
plot2 <- ggplot(nettle,
aes(x = MGS, y = Langs, label = Country)) +
geom_text()
# Plot double plot:

grid.arrange(plot1, plot2, ncol = 2)
```


```{r}
# Plotting pipeline with %>%:
nettle %>%
filter(MGS > 8) %>%
ggplot(aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```

```{r}
#how to inspect data frame:
str(nettle) # base R inspect the dataframe
glimpse(nettle) #tidyverse inspect the dataframe

```

```{r}
#median with $
median(nettle$Population) #in this case we need $ not comma as meadian takes one specific colum from the data frame
```

#Winter Ch 2.11.3-4


```{r}
head(nettle)
filter(nettle, Country == 'Benin')

filter(nettle, Country %in% c('Benin', 'Zaire')) #checks whether each value in a vector matches any value in a set of values — it returns TRUE/FALSE for each element

select(nettle, Langs)

filter(nettle, Country == 'Benin') %>% select(Langs)
```

```{r}
filter(nettle, Country == 'Benin') %>%
select(Population:MGS)

filter(nettle, Langs > 200)

filter(nettle, Langs > 200, Population < median(Population)) #Population < median(Population) means:Keep only countries whose population is below the median population of the whole dataset



```



```{r}
#Exercises 2.11.4
# Reduce the nettle tibble to small countries:
smallcountries <- filter(nettle, Population < 4)

# Create categorical MGS variable:
nettle_MGS <- mutate(smallcountries,
MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) #we can have string things together with the "&" like, to have 3 conditions

# Group tibble for later summarizing:
nettle_MGS_grouped <- group_by(nettle_MGS, MGS_cat)

# Compute language counts for categorical MGS variable:
summarize(nettle_MGS_grouped, LangSum = sum(Langs))
```


```{r}
# My try: Can you condense all of these steps
nettle_tibble <- filter(nettle, Population < 4) %>% mutate(MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) %>% group_by(MGS_cat) %>% summarize (LangSum = sum(Langs)) 
#mutate: i am gonna create sth new

nettle_tibble
```

##case_when
```{r}


nettle %>% 
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS < 4 ~ "very dry",
      MGS < 6 ~ "dry",
      MGS < 8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile") #para ordenar las categorias
  
  ) %>% 
  group_by(MGS_cat) %>% 
  summarize (LangSum = sum(Langs))
```


```{r} 
# get descriptive statistics for the smallpopulation data
nettle %>%
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS <  4 ~ "very dry",
      MGS <  6 ~ "dry",
      MGS <  8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile")
  ) %>%
  group_by(MGS_cat) %>%
  summarize(mean = mean(Langs) %>% round(2),
            sd = sd(Langs) %>% round(2),
            min = min(Langs),
            max = max(Langs))
```
#Winter Ch 3


```{r}
# Generate 50 random uniformly distributed numbers:
x <- runif(50)
# Check:
x
#para especificar rango de numeros
x <- runif(50, min = 2, max = 6)
head(x)
```

```{r}
#Plottng a histogram
hist(x, col = 'steelblue')
```

```{r}
#randmly distributed data
x <- rnorm(50)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2) #lty is type of line --dashed, lwd is width of line
```

```{r}
#specifying option arguments mean and sd (asignando valores a mean y sd):
x <- rnorm(50, mean = 5, sd = 2)
```

###mean and sd:
```{r}
mean(x)

```
```{r}
sd(x) 
```
##quantile 
```{r}
quantile(x) #(where a certain percentage of your data falls below a specific point)
```
quantile 16
```{r}
quantile(x, 0.16)
```
quantile 84
```{r}
quantile(x, 0.84)
```
sanity check
```{r}
mean(x)-sd(x)


```
```{r}
mean(x)+sd(x)


```
```{r}
# 2.5th percentile:
quantile(x, 0.025)
```
```{r}
# Should correspond to M – 2 * SD:
mean(x) - 2 * sd(x)
```
```{r}
# 97.5th percentile:
quantile(x, 0.975)
```
```{r}
# Should correspond to M + 2 * SD:
mean(x) + 2 * sd(x)
```
```{r}
#20 random numbers from a normal distribution (mean 0, SD 1)-not specified, see above
hist(rnorm(n = 20)) # execute repeatedly
```
## class activities

```{r}
# generate and plot a normal distribution
rnorm(50) %>%
  hist() # base r

# non-base r version
rnorm(50) %>%
  as.data.frame() %>%      # make a data frame
  rename(score = 1) %>%    # rename the variable
  ggplot(aes(x = score)) + # make the plot
  geom_histogram()

# set the mean and sd
rnorm(50, 71.82, 10.42) %>% #rnorm(number of values, mean, standard deviation)
  hist() 

# with ggplot
rnorm(50, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>% #cambia el nombre de la 1era columna a score
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# with 12 participants
rnorm(12, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")
# with 1200
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# how would we screen / filter scores > 100 and < 0
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# convert to sds
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  mutate(
    score.z = scale(score) # take the z score (standardize); z-score: How far a score is from the mean, measured in standard deviations.
  ) %>%
  ggplot(aes(x = score.z)) +
  geom_histogram(color = "black", fill = "white")
```
## CH 3 exercises
```{r}
war <- read_csv('warriner_2013_emotional_valence.csv')
war
```

```{r}
range(war$Val)
```
```{r}
filter(war, Val == min(Val) | Val == max(Val))
```
```{r}
filter(war, Val %in% range(Val))
```
```{r}
arrange(war, Val) # ascending order
```
```{r}
arrange(war, desc(Val)) # descending order
```
```{r}
mean(war$Val)
```
```{r}
sd(war$Val)
```
```{r}
mean(war$Val) - sd(war$Val)
```

```{r}
mean(war$Val) + sd(war$Val)
```
```{r}
quantile(war$Val, c(0.16, 0.84))
```
```{r}
median(war$Val)
```

```{r}
quantile(war$Val, 0.5)
```
Exercises Ch.3
With the Warriner et al. (2013) data, create a ggplot2 histogram and plot the mean
as a vertical line into the plot using geom_vline() and the xintercept aesthetic
(see Chapter 2). Can you additionally add vertical dashed lines to indicate where 68%
and 95% of the data lie? (Ignore any warning messages about binwidth that may arise).

```{r}
ggplot(war, aes(x = Val)) +
  geom_histogram(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```

Exercise 2:geo_histogram > geom_density
```{r}
ggplot(war, aes(x = Val)) +
  geom_density(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```
Exercise 3
Imagine reading a research paper about a grammaticality rating study. It is noted that
the mean acceptability rating for a particular grammatical construction is 5.25 with a
standard deviation of 0.4. Assuming normality, what is the interval within which you
expect 68% of the data to lie? What about 95% of the data? Do you think the assumption
of approximate normality is reasonable in this case?
```{r}
M <- 5.25
SD <- 0.4

lower_68 <- M - SD
upper_68 <- M + SD

lower_68
upper_68
```
95%

```{r}
lower_95 <- M - 2 * SD
upper_95 <- M + 2 * SD

lower_95
upper_95

```
problem: Likert-scale data is discrete (only whole numbers are possible responses), while the normal distribution is a continuous model (mismatch)
"While the calculated 68% and 95% intervals numerically fit within a plausible Likert scale (e.g., 1–7), there are still reasons to question the assumption of approximate normality. First, Likert-scale responses are discrete rather than continuous, which conflicts with the continuous nature that a normal distribution assumes. Second, since the mean (5.25) sits closer to the upper end of a typical scale, there is less room for values to spread out above the mean than below it, which could produce a skewed rather than symmetric distribution — undermining the normality assumption even though we cannot confirm this from the summary statistics alone."

variable based on low valence and high valence and then map into box plot


### boxplots

```{r}
# valence of words a vs p

war %>%
  mutate(
    word_type = case_when(
      str_detect(Word, "^a") ~ "a", #str_detect busca patterns within a text; ^ significa If it starts with a, put "a" in word_type
      str_detect(Word, "^p") ~ "p", #~"p" significa If the condition on the left is TRUE, put "p" in the new column
      .default = "error" #esta linea dice que hacer en caso de que no haya a ni p
    )
  ) %>%
  filter(word_type != "error") %>% #Keep only words that are not labeled "error"
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()

# creating a valence factor with levels low, medium, high
war %>%
  mutate(
    word_type = case_when(
      Val < 3 ~ "low",
      Val < 7 ~ "medium",
      Val >= 7 ~ "high",
      .default = "error"),
    word_type = fct_relevel(word_type, "low", "medium", "high")
  ) %>%
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()
```
#Winter CH4
```{r}
# Generate 50 random numbers:
x <- rnorm(50)
# Check first 6 values:
head(x)
# Create y's with intercept = 10 and slope = 3:
y <- 10 + 3 * x
plot(x, y, pch = 19)
#adding some error
error <- rnorm(50)
y <- 10 + 3 * x + error
plot(x, y, pch = 19)
```
```{r}
#linear model
xmdl <- lm(y ~ x)
xmdl
head(fitted(xmdl))
head(residuals(xmdl))
summary(xmdl)
```
```{r}
coef(xmdl)
# Alternative way of indexing (by name):
coef(xmdl)['(Intercept)']
#The following command computes the fitted value for an x-value of 10:
coef(xmdl)['(Intercept)'] + coef(xmdl)['x'] * 10
```
```{r}
#prediction for a sequence of numbers
xvals <- seq(from = -3, to = 3, by = 0.1) #from –3 to +3 using the seq() function. This particular sequence is specified to increase in 0.1 intervals (by = 0.1).
mypreds <- tibble(x = xvals) #The predict() function needs a data frame or tibble as input

mypreds$fit <- predict(xmdl, newdata = mypreds)
mypreds
```
##Linear model w/ tidyverse

```{r}
mydf <- tibble(x, y)
xmdl <- lm(y ~ x, data = mydf)
# Print tidy coefficient table to console:
tidy(xmdl)
# Extract estimate column from coefficient table:
tidy(xmdl)$estimate
# Check overall model performance:
glance(xmdl)
#plot
mydf %>% ggplot(aes(x = x, y = y)) +
geom_point() + geom_smooth(method = 'lm') +
theme_minimal()
```

#Winter Ch 5

```{r}
RTs <- c(600, 650, 700, 1000, 4000)
RTs
logRTs <- log(RTs)
logRTs
```

```{r}
exp(logRTs) # undo the logging
```

```{r}
ELP <- read_csv('ELP_frequency.csv')
ELP
```

```{r}
ELP <- mutate(ELP,
Log10Freq = log10(Freq),
LogRT = log(RT))
ELP
```

```{r}
#Raw frequencies
ELP %>% ggplot(aes(x = Freq, y = LogRT, label = Word)) +
geom_text() +
geom_smooth(method = 'lm') +
ggtitle('Log RT ~ raw frequency') +
theme_minimal()
```

```{r}
#log frequency values
ELP %>% ggplot(aes(x = Log10Freq, y = LogRT,
label = Word)) +
geom_text() +
geom_smooth(method = 'lm') +
ggtitle('Log RT ~ log frequency') +
theme_minimal()
```
```{r}
ELP_mdl <- lm(LogRT ~ Log10Freq, data = ELP)
tidy(ELP_mdl)
```

```{r}
b0 <- tidy(ELP_mdl)$estimate[1] # intercept
b1 <- tidy(ELP_mdl)$estimate[2] # slope
logRT_10freq <- b0 + b1 * 1
logRT_1000freq <- b0 + b1 * 3
logRT_10freq #predicted log response durations
logRT_1000freq
exp(logRT_10freq) #back to raw response durations, in miliseconds
exp(logRT_1000freq)
```
```{r}
ELP <- mutate(ELP,
Log10Freq_c = Log10Freq - mean(Log10Freq),
Log10Freq_z = Log10Freq_c / sd(Log10Freq_c)) #how many standard deviations a word's Log10Freq is from the mean
select(ELP, Freq, Log10Freq, Log10Freq_c, Log10Freq_z)
```

```{r}
# Same as before, different approach:
ELP <- mutate(ELP,
Log10Freq_c = scale(Log10Freq, scale = FALSE),
Log10Freq_z = scale(Log10Freq))
```

linear models with the centered and standardized
predictors:
```{r}

ELP_mdl_c <- lm(LogRT ~ Log10Freq_c, ELP) # centered
ELP_mdl_z <- lm(LogRT ~ Log10Freq_z, ELP) # z-scored
```
Comparing the coefficients from the different models:
```{r}
tidy(ELP_mdl) %>% select(term, estimate)
```
```{r}
tidy(ELP_mdl_c) %>% select(term, estimate) # x = 0 represents the mean of Log10Freq;the intercept (6.507966) is the predicted response at the mean; slope: a 1-unit increase in Log10Freq = 0.104-unit decrease in response
```

```{r}
tidy(ELP_mdl_z) %>% select(term, estimate) # Slope changes: 1 unit of Log10Freq_z = 1 SD of Log10Freq. For a 1 SD increase in Log10Freq, the predicted response decreases by 0.150103 units
```
```{r}
glance(ELP_mdl)$r.squared
glance(ELP_mdl_c)$r.squared
glance(ELP_mdl_z)$r.squared #r.squared: how much of the variation in Y is explained by the regression model
```

```{r}
glance(lm(LogRT ~ Freq, ELP))$r.squared # Nonlinear transformations of X can change R² and model fit
```
```{r}
with(ELP, cor(Log10Freq, LogRT)) #correlation is negative means that, as frequency increases, response durations become shorter.
```
## Recreates correlation using regression with z-scored X and Y; no intercept needed
```{r}
ELP_cor <- lm(scale(LogRT) ~ -1 + Log10Freq_z, ELP)
tidy(ELP_cor) %>% select(estimate) #explanation on p101
```
##exercise ch5-guesstimate

#Problem set 1
```{r}
data_stress <- read_csv("PS1/data_stress.csv")
data_answers <- read_csv("PS1/data_answers.csv")
glimpse(data_stress)

```
```{r}
glimpse(data_answers)
```
##no
```{r}
names(data_stress)
names(data_answers)
```
##data_stress response
```{r}
unique(data_stress$screen) #As each trial has several rows, "screen" column shows the label of each row, i.e., NA "instructions" "countdown"    "stimulus"     "response"    "timeout" 
  data_stress_clean <- filter(data_stress, screen == "response") #from screen column i keep only participants' responses
data_stress_clean 
```

```{r}
nrow(data_stress_clean) #the number of rows is the same as in the instructions
```
```{r}
length(unique(data_stress_clean$id)) #to chek the # of participants (38)
count(data_stress_clean, id) #all 38 participants have 80 trials
```

```{r}
count(data_stress_clean, type) #matches 38x48 (distractor) and 38x32(target)
```
##data_stress cleaning
In the data_stress 'el' has an accent, while in data_answer it does not. Also, what is ´eu´ in the answer sheet? typo?. These functions let me compare both data sets
```{r}
unique(data_stress_clean$response) #particiapnts response: "tu"       NA         "yo"       "él"       "ellos"    "nosotros"
unique(data_answers$answer) #answers: "nosotros" "yo"       "ellos"    "tu"       "el"       "eu" 
sum(is.na(data_stress_clean$response))  # to see # of trials with no response (86)
```
```{r}
filter(data_answers, answer == "eu") #so I am assuming there is a typo in the answer column and i will change eu into tu
```
##Fixing matches-response&answer
```{r}
data_stress_clean$response <- ifelse(data_stress_clean$response == "él", "el", data_stress_clean$response) #I am changing the el with accent with the el without accent and without modifying the rest
unique(data_stress_clean$response) #checking: "tu"       NA         "yo"       "el"       "ellos"    "nosotros"

sum(is.na(data_stress_clean$response)) #double check
```
```{r}
data_answers$answer <- ifelse(data_answers$answer == "eu", "tu", data_answers$answer)
unique(data_answers$answer)
nrow(data_answers) # 130
```
#I am trying to find a merge key:
```{r}
count(data_answers, sentence)
any(duplicated(data_answers$sentence)) # TRUE
data_answers[duplicated(data_answers$sentence), ]

```

```{r}
print(data_answers[duplicated(data_answers$sentence), c("sentence", "verb conjugation", "answer")], width = Inf)
```
```{r}
subset(data_answers, sentence == "Aca lo invito")
```

```{r}
head(data_stress_clean)
head(data_answers)
unique(data_answers$answer)
```


```{r}
data_stress_clean
head(data_stress_clean[, c("verb", "tense", "audio")], 12)
unique(data_stress_clean$audio)
```


```{r}
head(unique(data_stress_clean$verb), 20)
head(unique(data_answers$`verb conjugation`), 20)
```
##filenames extraction
Extracting sentences from audio column in data_stress
```{r}
data_stress_clean <- data_stress_clean %>%
  mutate(sentence_clean = gsub(".*_", "", audio)) %>% # keep only what comes after the last underscore
  mutate(sentence_clean = gsub(".mp3", "", sentence_clean)) # remove the ".mp3"
head(data_stress_clean$sentence_clean)
```
```{r}
setdiff(data_stress_clean$sentence_clean, data_answers$sentence) #character(0)
```
```{r}
any(duplicated(paste(data_answers$sentence, data_answers$tense))) #paste() glues the two columns into one text per row, so this asks, "does any sentence + tense combination appear more than once?"
```
```{r}
unique(data_stress_clean$tense)
unique(data_answers$tense) #tenses are written similarly in both files
```
```{r}
head(data_stress_clean)
head(data_answers_clean)
```
##merged
```{r}
data_answers_clean <- data_answers[, c("sentence", "tense", "answer")]
data_stress_merged <- merge(data_stress_clean, data_answers_clean,
                            by.x = c("sentence_clean", "tense"),
                            by.y = c("sentence", "tense")
                            )

```
##scored

```{r}
data_stress_scored <- data_stress_merged %>%
  mutate(score = ifelse(response == answer, 1, 0))

data_stress_scored %>%
  select(response, answer, score) %>%
  head(10)
```


```{r}
head(data_stress_scored)
```
##mean participant/type
```{r}

data_stress_means <- data_stress_scored %>%
  group_by(id, type) %>% #each combination of participant (id) and type is a  group
  summarise(mean_score = mean(score, na.rm = TRUE)) #mean_score is a new colum w/ each group's mean
head(data_stress_means)
unique(data_stress_scored$response) #this should always keep el, tu without the accent
```
##data proficiency
```{r}
data_proficiency <- read_csv("PS1/data_proficiency.csv")
glimpse(data_proficiency)
names(data_proficiency) # [1] "id"  "trial"   "response"  [4] "onset_time"             "clock_time"             "reaction_time"         [7] "absolute_onset_time"    "absolute_clock_time"    "absolute_reaction_time" [10] "tag"                    "answer"   
dim(data_proficiency) #[1] 3188   11
```
```{r}
unique(data_proficiency$tag) #tag tiene NA        "rawr"    "click"   "timeout"
data_proficiency %>%
  select(id, trial, response, tag, answer) %>%
  head(20)

```
```{r}
count(data_proficiency, tag) #dentro de column tag, click parece ser la respuesta del participant
count(data_proficiency, id) #la mayoria tiene 85 trials, excepto 2
data_proficiency_counts <- count(data_proficiency, id)
subset(data_proficiency_counts, n != 85) #sale los datos de los 2 que no completaron todos los ejercicios. Decidi no incluirlos
```
```{r}
subset(data_proficiency, id == 15423803) #BORRAR
unique(data_proficiency$tag)
```


```{r}
table(data_proficiency$tag, is.na(data_proficiency$response)) #is.na asks is `response' missing, FALSE=not missing
head(subset(data_proficiency, tag == "click"), 10) #this shows that the tag "click" has participants' answers
head(subset(data_proficiency, tag == "rawr"), 10) #response have NA
head(subset(data_proficiency, tag == "timeout"), 10) #response have literally "timeout"
```
###timeout into NA
```{r}
data_proficiency_clean <- data_proficiency %>%
  filter(tag == "click" | tag == "timeout") %>%
  mutate(response = ifelse(response == "timeout", NA, response))
unique(data_proficiency_clean$response) #ahora los "timeout" cambiaron a NA en al columna response
nrow(data_proficiency_clean) #1862, the same as above with table(it was just to check)
sum(is.na(data_proficiency_clean$response)) #to check if "timeout" change into NA
```
```{r}
data_proficiency_clean <- data_proficiency_clean %>%
  select(id, trial, response, answer)
head(data_proficiency_clean)
data_proficiency_scored <- data_proficiency_clean %>%
  mutate(score = ifelse(response == answer, 1, 0))
table(data_proficiency_scored$score, useNA = "ifany")
```

```{r}
head(subset(data_proficiency_scored, score == 0, select = c(response, answer)), 20)
```
```{r}
setdiff(data_proficiency_scored$response, data_proficiency_scored$answer)
```
```{r}
setdiff(data_proficiency_scored$answer, data_proficiency_scored$response) #this shows that all answers where chosen at least once
head(data_proficiency_scored)

data_proficiency_means <- data_proficiency_scored %>%
  group_by(id) %>% #solo id porque habian mas de una columna por id y ahora solo 1
  summarise(proficiency = mean(score, na.rm = TRUE))
nrow(data_proficiency_means) #38 en total, that is, #of participants
range(data_proficiency_means$proficiency)
head(data_proficiency_means)
```
```{r}
count(data_proficiency_clean, id) %>% filter(n != 50)
incomplete_ids <- c(15411133, 15423803)
incomplete_ids
data_proficiency_means <- data_proficiency_means %>%
  filter(!(id %in% incomplete_ids)) #significa: manten todos exepto los de arriba
nrow(data_proficiency_means) #36, ya excluidos los dos de arriba
```
##incomplete_ids
```{r}
data_stress_means <- data_stress_means %>%
  filter(!(id %in% incomplete_ids))
```
###ungrouping incomplete ids
```{r}
#ungrouping the two participants who did not finish the task
data_stress_means <- data_stress_means %>%
  ungroup() %>%
  filter(!(id %in% incomplete_ids))
nrow(data_stress_means) #72, that is, 36 participants x 2types
```
##1 row instead of 2
NO LO USE
data_stress_wide <- data_stress_means %>%
  pivot_wider(names_from = type, values_from = mean_score) #makes one extra column instead of 2 rows
head(data_stress_wide)
nrow(data_stress_wide)
NO LO USE
data_final <- merge(data_stress_wide, data_proficiency_means, by = "id")
head(data_final)
nrow(data_final)
sum(is.na(data_final)) #sanity check, debe ser 0

```{r}
data_long <- merge(data_stress_means, data_proficiency_means, by = "id")
nrow(data_long)
head(data_long)
```
##plot score_type
```{r}
ggplot(data_long, aes(x = type, y = mean_score)) +
  geom_boxplot()
```

The boxplot shows that the median of participants' mean scores is generally higher for distractor items than for target items. The median score is approximately 0.80 for distractors and 0.60 for targets. We can also see that the distributions overlap. This may indicate that there is variation in scores and that the difference between item types is not clear-cut.

##plot_prof and score
```{r}

data_long %>%
  ggplot(aes(x = proficiency, y = mean_score, color=type)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_minimal()
```

As proficiency increases, mean scores increase for both distractor and target items. The distractor line sits above the target line across the range of proficiency, and the lines are roughly parallel. The points are loosely clustered around the fitted lines, which indicates that there is reasonable variability between participants with similar proficiency. The shaded bands overlap slightly at the extremes.

##linear model
```{r}
ps_model <- lm(mean_score ~ proficiency + type, data = data_long)
ps_model
tidy(ps_model)
summary(ps_model)
```

##Data analysis


**What can we say about the relationship between our predictors and score?** 

The model suggests that both proficiency and type are related to mean score, but in different ways. Proficiency has a positive coefficient of 0.63174, which means that for every one-unit increase in proficiency, the predicted mean score increases by 0.63174. Therefore, participants with higher proficiency are predicted to have higher mean scores. 

Type is also related to mean score. Since distractor is the reference category, the typetarget coefficient of −0.17614 means that target items have a predicted mean score that is 0.17614 lower than distractor items. Thus, the model predicts lower scores for target items than for distractor items at the same level of proficiency.

**As part of your interpretation, explain what the intercept represents in this model. Is it a meaningful value, given the range of proficiency scores in your data?**

The intercept (0.41153) represents the predicted mean score when proficiency is zero and the item type is distractor, the reference category. However, as the proficiency=0 does nto occur in our data-- the range of proficiency is 0.140000  to 0.877551-- tthe intercept is not particularly meaningful for interpreting the observed data; probably, centering the data would be more useful.

**Finally, note that each participant contributes two mean scores to this model (one per item type). Briefly explain why that might be a problem for a linear model. We will return to this issue when we cover mixed-effects models.**
Each participant contributes two observations (mean scores for each type), and those observations may be correlated because they come from the same participant. Therefore, the model may not take this relationship into account and may treat the two observations as independent of one another