---
title: "CLASSCODE"
author: "JEFFERSON SARANSIG"
date: "2026-08-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("janitor")
library(tidyverse)
library(gridExtra)
library(janitor)
```

# Winter CH 1; code from ch 1 of winter textbook
```{r}
numbers <- c(18, 200, 92, 14)
sum(numbers)
mean(numbers)
sd(numbers) %>% round(digits = 2) #the "digits = " se usa solo para recordarnos, en realidad se puede hacer solo round(2). "%>%" help us string operations
names <- c("a", "b", "c")
names

```


```{r}
list.files()
head(nettle) #head() shows the first few rows of a data frame
nettle <- read.csv('nettle_1999_climate.csv')
head(nettle)
tail(nettle) #head() shows the last few rows of a data frame
```


##histogram (langs+color)
```{r} 

hist(nettle$Langs) # the $ operator extracts a specific column (or list element) by name from a data frame or list.
hist(nettle$Langs, col = 'salmon')
```


```{r}
hist(nettle$Langs, col = 'steelblue')
```

```{r}
x_values <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
mean_x <- mean(x_values)
mean_x #para sacar el valor de la media
```

```{r}
x <- c(2, 3, 4, '4')
x <- as.numeric(x)
mean(x)

```

```{r}
head(nettle)
nettle[2,5]
nettle[1:4, 1:2]
nettle[nettle$Country == 'Bangladesh', 'Langs']
nettle[1:4, ]$Langs[2]
nettle[1:4, c('Country', 'Langs')]
head(nettle[,])

```

## read.csv function
```{r}
# Load data:
nettle <- read.csv('nettle_1999_climate.csv') #"read.csv" is part of Base R

# Convert data frame to tibble:
nettle <- as_tibble(nettle)
nettle
```
##read_csv function
```{r}
nettle <- read_csv('nettle_1999_climate.csv')
nettle

```

**filter con piping**

```{r} 
nettle%>%filter(Area > 6)
```
**ggplot con points**
```{r}
ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs)) # "mapping" can be omitted
```

```{r}
ggplot(nettle, aes(x = Population, y = Area)) +
  geom_point()
```



**ggplot con texto**
```{r}
ggplot(nettle, aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```


```{r}
ggsave('nettle.png', width = 8, height = 6) #to save an image
```

```{r}
# Create plots and save them in plot1 and plot2:
plot1 <- ggplot(nettle) +
geom_point(mapping = aes(x = MGS, y = Langs))
plot2 <- ggplot(nettle,
aes(x = MGS, y = Langs, label = Country)) +
geom_text()
# Plot double plot:
library(gridExtra)
grid.arrange(plot1, plot2, ncol = 2)
```


```{r}
# Plotting pipeline with %>%:
nettle %>%
filter(MGS > 8) %>%
ggplot(aes(x = MGS, y = Langs, label = Country)) +
geom_text()
```

```{r}
#how to inspect data frame:
str(nettle) # base R inspect the dataframe
glimpse(nettle) #tidyverse inspect the dataframe

```

```{r}
#median with $
median(nettle$Population) #in this case we need $ not comma as meadian takes one specific colum from the data frame
```

#Winter Ch 2.11.3-4


```{r}
head(nettle)
filter(nettle, Country == 'Benin')

filter(nettle, Country %in% c('Benin', 'Zaire')) #checks whether each value in a vector matches any value in a set of values — it returns TRUE/FALSE for each element

select(nettle, Langs)

filter(nettle, Country == 'Benin') %>% select(Langs)
```

```{r}
filter(nettle, Country == 'Benin') %>%
select(Population:MGS)

filter(nettle, Langs > 200)

filter(nettle, Langs > 200, Population < median(Population)) #Population < median(Population) means:Keep only countries whose population is below the median population of the whole dataset



```



```{r}
#Exercises 2.11.4
# Reduce the nettle tibble to small countries:
smallcountries <- filter(nettle, Population < 4)

# Create categorical MGS variable:
nettle_MGS <- mutate(smallcountries,
MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) #we can have string things together with the "&" like, to have 3 conditions

# Group tibble for later summarizing:
nettle_MGS_grouped <- group_by(nettle_MGS, MGS_cat)

# Compute language counts for categorical MGS variable:
summarize(nettle_MGS_grouped, LangSum = sum(Langs))
```


```{r}
# My try: Can you condense all of these steps
nettle_tibble <- filter(nettle, Population < 4) %>% mutate(MGS_cat = ifelse(MGS < 6, 'dry', 'fertile')) %>% group_by(MGS_cat) %>% summarize (LangSum = sum(Langs)) 
#mutate: i am gonna create sth new

nettle_tibble
```

##case_when
```{r}


nettle %>% 
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS < 4 ~ "very dry",
      MGS < 6 ~ "dry",
      MGS < 8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile") #para ordenar las categorias
  
  ) %>% 
  group_by(MGS_cat) %>% 
  summarize (LangSum = sum(Langs))
```


```{r} 
# get descriptive statistics for the smallpopulation data
nettle %>%
  filter(Population < 4) %>%
  mutate(
    MGS_cat = case_when(
      MGS <  4 ~ "very dry",
      MGS <  6 ~ "dry",
      MGS <  8 ~ "fertile",
      MGS >= 8 ~ "very fertile",
      .unmatched = "error"
    ),
    MGS_cat = fct_relevel(MGS_cat, "very dry", "dry", "fertile", "very fertile")
  ) %>%
  group_by(MGS_cat) %>%
  summarize(mean = mean(Langs) %>% round(2),
            sd = sd(Langs) %>% round(2),
            min = min(Langs),
            max = max(Langs))
```
#Winter Ch 3


```{r}
# Generate 50 random uniformly distributed numbers:
x <- runif(50)
# Check:
x
#para especificar rango de numeros
x <- runif(50, min = 2, max = 6)
head(x)
```

```{r}
#Plottng a histogram
hist(x, col = 'steelblue')
```

```{r}
#randmly distributed data
x <- rnorm(50)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2) #lty is type of line --dashed, lwd is width of line
```

```{r}
#specifying option arguments mean and sd (asignando valores a mean y sd):
x <- rnorm(50, mean = 5, sd = 2)
```

###mean and sd:
```{r}
mean(x)

```
```{r}
sd(x) 
```
##quantile 
```{r}
quantile(x) #(where a certain percentage of your data falls below a specific point)
```
quantile 16
```{r}
quantile(x, 0.16)
```
quantile 84
```{r}
quantile(x, 0.84)
```
sanity check
```{r}
mean(x)-sd(x)


```
```{r}
mean(x)+sd(x)


```
```{r}
# 2.5th percentile:
quantile(x, 0.025)
```
```{r}
# Should correspond to M – 2 * SD:
mean(x) - 2 * sd(x)
```
```{r}
# 97.5th percentile:
quantile(x, 0.975)
```
```{r}
# Should correspond to M + 2 * SD:
mean(x) + 2 * sd(x)
```
```{r}
#20 random numbers from a normal distribution (mean 0, SD 1)-not specified, see above
hist(rnorm(n = 20)) # execute repeatedly
```
## class activities

```{r}
# generate and plot a normal distribution
rnorm(50) %>%
  hist() # base r

# non-base r version
rnorm(50) %>%
  as.data.frame() %>%      # make a data frame
  rename(score = 1) %>%    # rename the variable
  ggplot(aes(x = score)) + # make the plot
  geom_histogram()

# set the mean and sd
rnorm(50, 71.82, 10.42) %>% #rnorm(number of values, mean, standard deviation)
  hist() 

# with ggplot
rnorm(50, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>% #cambia el nombre de la 1era columna a score
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# with 12 participants
rnorm(12, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")
# with 1200
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# how would we screen / filter scores > 100 and < 0
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  ggplot(aes(x = score)) +
  geom_histogram(color = "black", fill = "white")

# convert to sds
rnorm(1200, 71.82, 10.42) %>%
  as.data.frame() %>%
  rename(score = 1) %>%
  filter(score <= 100) %>%
  mutate(
    score.z = scale(score) # take the z score (standardize); z-score: How far a score is from the mean, measured in standard deviations.
  ) %>%
  ggplot(aes(x = score.z)) +
  geom_histogram(color = "black", fill = "white")
```
## CH 3 exercises
```{r}
war <- read_csv('warriner_2013_emotional_valence.csv')
war
```

```{r}
range(war$Val)
```
```{r}
filter(war, Val == min(Val) | Val == max(Val))
```
```{r}
filter(war, Val %in% range(Val))
```
```{r}
arrange(war, Val) # ascending order
```
```{r}
arrange(war, desc(Val)) # descending order
```
```{r}
mean(war$Val)
```
```{r}
sd(war$Val)
```
```{r}
mean(war$Val) - sd(war$Val)
```

```{r}
mean(war$Val) + sd(war$Val)
```
```{r}
quantile(war$Val, c(0.16, 0.84))
```
```{r}
median(war$Val)
```

```{r}
quantile(war$Val, 0.5)
```
Exercises Ch.3
With the Warriner et al. (2013) data, create a ggplot2 histogram and plot the mean
as a vertical line into the plot using geom_vline() and the xintercept aesthetic
(see Chapter 2). Can you additionally add vertical dashed lines to indicate where 68%
and 95% of the data lie? (Ignore any warning messages about binwidth that may arise).

```{r}
ggplot(war, aes(x = Val)) +
  geom_histogram(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```

Exercise 2:geo_histogram > geom_density
```{r}
ggplot(war, aes(x = Val)) +
  geom_density(fill = 'peachpuff3') +
  geom_vline(aes(xintercept = mean(war$Val)), linetype = 2) +
  geom_vline(xintercept = quantile(war$Val, c(0.16, 0.84)), linetype = 3) +
  geom_vline(xintercept = quantile(war$Val, c(0.025, 0.975)), linetype = 4) +
  theme_minimal()
```
Exercise 3
Imagine reading a research paper about a grammaticality rating study. It is noted that
the mean acceptability rating for a particular grammatical construction is 5.25 with a
standard deviation of 0.4. Assuming normality, what is the interval within which you
expect 68% of the data to lie? What about 95% of the data? Do you think the assumption
of approximate normality is reasonable in this case?
```{r}
M <- 5.25
SD <- 0.4

lower_68 <- M - SD
upper_68 <- M + SD

lower_68
upper_68
```
95%

```{r}
lower_95 <- M - 2 * SD
upper_95 <- M + 2 * SD

lower_95
upper_95

```
problem: Likert-scale data is discrete (only whole numbers are possible responses), while the normal distribution is a continuous model (mismatch)
"While the calculated 68% and 95% intervals numerically fit within a plausible Likert scale (e.g., 1–7), there are still reasons to question the assumption of approximate normality. First, Likert-scale responses are discrete rather than continuous, which conflicts with the continuous nature that a normal distribution assumes. Second, since the mean (5.25) sits closer to the upper end of a typical scale, there is less room for values to spread out above the mean than below it, which could produce a skewed rather than symmetric distribution — undermining the normality assumption even though we cannot confirm this from the summary statistics alone."

variable based on low valence and high valence and then map into box plot


### boxplots

```{r}
# valence of words a vs p
library(stringr) #is a package for working with text/strings
war %>%
  mutate(
    word_type = case_when(
      str_detect(Word, "^a") ~ "a", #str_detect busca patterns within a text; ^ significa If it starts with a, put "a" in word_type
      str_detect(Word, "^p") ~ "p", #~"p" significa If the condition on the left is TRUE, put "p" in the new column
      .default = "error" #esta linea dice que hacer en caso de que no haya a ni p
    )
  ) %>%
  filter(word_type != "error") %>% #Keep only words that are not labeled "error"
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()

# creating a valence factor with levels low, medium, high
war %>%
  mutate(
    word_type = case_when(
      Val < 3 ~ "low",
      Val < 7 ~ "medium",
      Val >= 7 ~ "high",
      .default = "error"),
    word_type = fct_relevel(word_type, "low", "medium", "high")
  ) %>%
  ggplot(aes(x = word_type, y = Val)) +
  geom_boxplot() +
  theme_minimal()
```


